---
title: "Big Data Hw1"
author: "EECS 107 0310003 Hung I Chiu"
date: "2017/3/16"
output: pdf_document
---

```{r setup, include=FALSE, cache = TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(plotly)
```

To TA:

This report is made by R markdown, but since I had trouble with typing Chinese characters in R markdown, so I decided to hand in my report in an English version.

Besides the data of December 2016 reaches almost 1 GB, my computer isn't good enough to handle much bigger dataset with R Language. I only analyze the data of December, 2016. Also, I think that the mean and tools to analyze won't varied even the dataset is Green or HLV or other months. 

First of all, the folder is not a standard CSV extension file on account of the EOL -- end of line -- character. The standard CSV file has "\\n" as the EOL, but as for the dataset, it was a double comma such as ",,". Consequently, I have to open the file by vim and delete all the double commas before I read the data by R console.

- pu_hour stands for pickup hour
- do_hour stands for dropoff hour
- distance stands for long or chort distance. I set the boundary at 15km because of my own experience.
```{r read, include=TRUE, cache = TRUE}
setwd('~/Downloads/')
data <- tbl_df(read.table(file="./yellow.csv", header=TRUE, sep=","))

data <- data %>%
  mutate(
    pu_hour = substr(as.character(tpep_pickup_datetime), 12, 13),
    do_hour = substr(as.character(tpep_pickup_datetime), 12, 13),
    distance = ifelse(trip_distance > 15, 1, 0)
    )

data %>% head(10)
```

### Q1
What regions have most pickups and drop-offs?

#### Pickup Region vs Count
We can see that LocationID of 237 has the most pickup.
```{r Q2-PU, include=TRUE, cache = TRUE}
data %>%
  group_by(PULocationID) %>%
  summarise(n = n()) %>%
  arrange(desc(n))
```

#### Dropoff Region vs Count
We can see that LocationID of 236 has the most dropoff.
```{r Q2-DO, include=TRUE, cache = TRUE}
data %>%
  group_by(DOLocationID) %>%
  summarise(n = n()) %>%
  arrange(desc(n))
```

#### (Pickup, Dropoff) Region vs Count
According to the above and the below table, we can infer that 236 and 237 might be neighbors.
```{r Q2-PU-DO, include=TRUE, cache = TRUE}
data %>%
  group_by(PULocationID, DOLocationID) %>%
  summarise(n = n()) %>%
  arrange(desc(n))
```

\newpage

### Q2
When are the peak hours and off-peak hours in taking taxi?

#### Pickup Hour vs Count
The peak hours would be 18 ~ 19 o'clock, and the off-peak hours would be 0 ~ 7 o'clock
```{r Q2-1, include=TRUE, cache = TRUE}
data %>%
  ggplot(aes(x=pu_hour)) +
  geom_bar(fill = "red", alpha = 0.2) +
  ggtitle("pickup hour vs count") +
  labs(x="hour", y="n")
```

\newpage

#### Dropoff Hour vs Count

Same as above.
```{r Q2-2, include=TRUE, cache = TRUE}
data %>%
  ggplot(aes(x=do_hour)) +
  geom_bar(fill = "red", alpha = 0.2) +
  ggtitle("dropoff hour vs count") +
  labs(x="hour", y="n")
```

\newpage

#### Q3
What differences exist between short and long distance trips of taking taxi?

I summarise columns that might have relation with distance.

distance with 0 means <15km and with 1 means > 15km.
```{r Q3-1, include=TRUE, cache = TRUE}
data %>%
  group_by(distance) %>%
  summarise(
    n= n(),
    avg_passenger = mean(passenger_count),
    avg_fare_amount = mean(fare_amount),
    avg_total_amount = mean(total_amount)
  )
```
\newpage

Later, since average couldn't show the distribution of the data point, I make scatter plots of passenger_count, fare_count, total_amount.

```{r Q3-2, include=TRUE, cache = TRUE}
data %>%
  sample_n(100000) %>%
  ggplot(aes(x=as.factor(distance), y=passenger_count)) +
  geom_point(alpha=0.2, size=1, color="red") +  
  ggtitle("distance vs passenger_count") +
  labs(x="distance", y="passenger_count")
```

\newpage

```{r Q3-2-2, include=TRUE, cache = TRUE}
data %>%
  sample_n(100000) %>%
  ggplot(aes(x=as.factor(distance), y=fare_amount)) +
  geom_point(alpha=0.2, size=1, color="red") +  
  ggtitle("distance vs fare_amount") +
  labs(x="distance", y="fare_amount")
```

\newpage

```{r Q3-2-3, include=TRUE, cache = TRUE}
data %>%
  sample_n(100000) %>%
  ggplot(aes(x=as.factor(distance), y=total_amount)) +
  geom_point(alpha=0.2, size=1, color="red") +  
  ggtitle("distance vs total_amount") +
  labs(x="distance", y="total_amount")
```

\newpage

It's worth noting that the unit of y axis is percentage, not count. 
```{r Q3-3, include=TRUE, cache = TRUE}
data %>%
  filter(distance == 0) %>%
  ggplot(aes(x=payment_type)) +
  geom_histogram(
    aes(y=(..count..)/sum(..count..) * 100), bins=50, fill = "red", alpha = 0.2) +
  ggtitle("payment type vs percentage (<=15km)") +
  labs(x="payment type", y="percentage (%)")
```

\newpage

```{r Q3-3-2, include=TRUE, cache = TRUE}
data %>%
  filter(distance == 1) %>%
  ggplot(aes(x=payment_type)) +
  geom_histogram(
    aes(y=(..count..)/sum(..count..) * 100), bins=50, fill = "red", alpha = 0.2) +
  ggtitle("payment type vs percentage (>15km)") +
  labs(x="payment type", y="percentage (%)")
```

\newpage

We can see that the pickup hour varies between long and short distance.
```{r Q3-4, include=TRUE, cache = TRUE}
data %>%
  filter(distance == 0) %>%
  ggplot(aes(x=factor(pu_hour))) +
  geom_bar(aes(y=(..count..)/sum(..count..) * 100), fill = "red", alpha = 0.2) +
  ggtitle("pickup hour vs percentage (<=15km)") +
  facet_wrap("distance") +
  labs(x="pickup hour", y="percentage")
```

\newpage

```{r Q3-4-2, include=TRUE, cache = TRUE}
data %>%
  filter(distance == 1) %>%
  ggplot(aes(x=factor(pu_hour))) +
  geom_bar(aes(y=(..count..)/sum(..count..) * 100), fill = "red", alpha = 0.2) +
  ggtitle("pickup hour vs percentage (>15km)") +
  facet_wrap("distance") +
  labs(x="pickup hour", y="percentage")
```
